---
title: "The relationship between Internet Use and GDP impact through Climate Change"
author: "Tejveer Rai"
output: html_document

---

```{r setup, include=FALSE}
library("ggplot2")
library("dplyr")
library("knitr")
library("tidyr")
library("maps")
library("ggvis")
library("plotly")

gdp_with_climate <- read.csv("data/GDPcap_ClimateChange_RCP85_SSP5.csv", stringsAsFactors = FALSE)
gdp_without_climate <- read.csv("data/GDPcap_NOClimateChange_RCP85_SSP5.csv", stringsAsFactors = FALSE)
worldbank_country <- read.csv("data/WorldBank_CountryIndicators.csv", stringsAsFactors = FALSE)

CalcPercent<-function(x,y){ #calculate the percent difference
  return((x - y)/ y)
}

gdp_comparison <- left_join(gdp_with_climate, gdp_without_climate, by = c("ISO3" = "ISO3")) %>% select(ISO3, name.x, X2010.x, X2010.y, X2030.x, X2030.y, X2050.x, X2050.y, X2070.x, X2070.y, X2090.x, X2090.y) %>% mutate(diff_2010 = CalcPercent(X2010.x, X2010.y), diff_2030 = CalcPercent(X2030.x, X2030.y), diff_2050 = CalcPercent(X2050.x, X2050.y), diff_2070 = CalcPercent(X2070.x, X2070.y), diff_2090 = CalcPercent(X2090.x, X2090.y)) %>% select(ISO3,name.x, diff_2010, diff_2030, diff_2050, diff_2070, diff_2090) 

colnames(gdp_comparison)
internet_users<- filter(worldbank_country, Series.Code == "IT.NET.USER.P2")

combined<-full_join(gdp_comparison, internet_users, by = c("ISO3" = "Country.Code"))

populations<- filter(worldbank_country, Series.Code == "SP.POP.TOTL")

```
##Introduction

Climate change has been a trending debate topic all around the world in the past years. It's clear that there could be an enivornmental impact on our planet from climate change, but there may also be a significant _economic_ impact as well! Through the data collected by **Burke, Hsiang, and Miguel (2015)**, we can analyze the difference in a country's GDP with and without climate change. The report below processes their data in 20-year time intervals, from 2010 to 2090. In addition, data collected by the **world bank** provides information on internet users for each country. They have recorded how many people for every 100 citizens use the internet in a country. By analyzing both of these sets of data together, it's clear that the majority of countries with a high number of internet users saw a positive percentage change from a GDP _without_ climate change compared to a GDP _with_ climate change. 

##Summary

### GDP & Climate Change

Using the data collected by the report, we can analyze the relationship between climate change and a country's GDP. In the following data visuals, we will be able to see how an increase in climate change while time passes correlates to an increase or decrease to a GDP. The data begins in 2010, where every country starts at 0% change. 

This graph shows the difference between a GDP with climate change and a GDP without climate change after 20 years has passed. The data is for the year 2030 and the GDP difference is ranked from -3% to 8.5%. We can see that the difference is fairly minimal at this stage. Almost all countries show little change in GDP when comparing climate change to no climate change. 

```{r 2030 graph, fig.width=25, echo=FALSE}

ggplot(data = gdp_comparison) + 
  geom_bar(mapping = aes(x = ISO3, y = diff_2030), fill = "black", stat="identity") +
  ylim(-3, 8.5) + theme(axis.title = element_text(size=30), plot.title = element_text(size=30), axis.text.x = element_text(angle=90, hjust=1)) + ggtitle("2030 Difference in GDP") + xlab("Country Code") + ylab("% Change") 

```

Another 20 years has passed and we can see the how nearly all countries have faced a larger percentage change. Comparing this to the 2030 data, we can confidently infer that climate change causes a larger impact on a country's GDP in 2050. At this stage, there is a higher degree of increase or decrease from the 2030 data but there is not many extreme values. Although, we can see that the majority of the countries are facing a negative GDP difference as a result of climate change.

```{r 2050 graph, fig.width=25, echo=FALSE}
ggplot(data = gdp_comparison) + 
  geom_bar(mapping = aes(x = ISO3, y = diff_2050), fill = "black", stat="identity") +
  ylim(-3, 8.5) + theme(axis.title = element_text(size=30), plot.title = element_text(size=30), axis.text.x = element_text(angle=90, hjust=1)) + ggtitle("2050 Difference in GDP") + xlab("Country Code") + ylab("% Change")
```

Here we have the visual for the year 2090. The change is major from 2050 and even more drastic from 2030. It's clear that the impact climate change is having on every country's GDP grows as time passes. A large majority of the countries have seen a negative difference at this point. Looking at the positive side, Mongolia possess an extreme value of a positive difference of 8.3%, almost 5% above the next highest! 

```{r 2090 graph, fig.width=25, echo = FALSE}
ggplot(data = gdp_comparison) + 
  geom_bar(mapping = aes(x = ISO3, y = diff_2090), fill = "black", stat="identity") +
  ylim(-3, 8.5) + theme(axis.title = element_text(size=30), plot.title = element_text(size=30), axis.text.x = element_text(angle=90, hjust=1)) + ggtitle("2090 Difference in GDP") + xlab("Country Code") + ylab("% Change")
```

### Internet Users

Now, we can see how the number of internet users in a country plays a roll in the overall difference of GDP. The graph below shows the GDP difference in climate change vs. no climate change in 2090. Although, each bar is color coded to represent the number of internet users. The pallet indicates blue as the lowest amount of internet users per 100 citizens, and red represents the highest amount. An analysis of the graph shows that the majority of countries who saw a positive difference had a very high number of internet users, with the exception of Mongolia. Almost all of the countries who have a negative percentage difference are colored in a shade of blue, meaning they have a low number of internet users in their country.

```{r internet graph, fig.width=25, fig.height=10, echo = FALSE}

  ggplot(data = combined) + 
  geom_bar(mapping = aes(x = ISO3, y = diff_2090, color = value), fill = "white", stat="identity")+ scale_colour_gradient(limits=c(0, 100), low="blue", high="red") + theme(axis.text.x = element_text(angle=90, hjust=1)) + ylim(-3, 8.5) + ggtitle("2090 Difference in GDP vs. Number of Internet Users") + xlab("Country Code") + ylab("% Change") + labs(color="Number of Internet Users per 100: ") + theme(axis.title = element_text(size=30), plot.title = element_text(size=30),legend.position = "bottom", legend.text=element_text(size=25), legend.title = element_text(size=30), legend.key.height = unit(4, 'lines'), legend.key.width = unit(5, 'lines'))

```

## Map Visualization

We can get a better understanding of how the differences in GDP are distributed by using a visual map. The four maps are color coded to represent the percentage change from a GDP without climate change to a GDP with climate change for each country. The included visual represents the data collected for 2030, 2050, 2070 and 2090. It's clear to see that the differences grow in severity as time progresses. 2030 has an almost constant color across the map while 2090 shows a wide variety of colors. This a good representation of how the impact on GDP by climate change grows as time goes on. 

```{r world map, fig.width=25, fig.height = 15, echo = FALSE}
world <- map_data("world")
world<- mutate(world, ISO3 = iso.alpha(region, n = 3))
world_gdp_combined<- full_join(world, gdp_comparison, by = c("ISO3" = "ISO3"))

world_gdp_combined<- select(world_gdp_combined, long, lat, group, order, region, ISO3, 
                            diff_2030, diff_2050, diff_2070, diff_2090)

world_gdp_combined<- gather(world_gdp_combined, 
                        key = years, 
                        value = difference, 
                        diff_2030, diff_2050, 
                        diff_2070, diff_2090
                       )

world_gdp_combined$difference<-cut(world_gdp_combined$difference, breaks=c(-1, -.6, -.2, .2, .6, 1, 3, 9), 
                                   labels = c("-1 to -0.6", "-0.6 to -0.2", "-0.2 to 0.2", "0.2 to 0.6","0.6 to 1", "1 to 3", 
                                              "3 to 9"))

world_gdp_combined<-filter(world_gdp_combined, difference != is.na(difference))

labels<- c(diff_2030 = "2030", diff_2050 = "2050", diff_2070 = "2070", diff_2090 = "2090")

ggplot(data = world_gdp_combined) +
 geom_polygon(aes(x = long, y = lat, group = group, fill = difference)) +
 coord_quickmap() + facet_wrap(~years, labeller=labeller(years = labels)) + labs(fill="Difference in GDP (%) from presence of climate change: ") + theme(legend.position = "bottom", plot.title = element_text(size=30), axis.text = element_text(size=20), axis.title = element_text(size=30), legend.text=element_text(size=25), legend.title = element_text(size=30), legend.key.height = unit(4, 'lines'), strip.text.x = element_text(size = 25)) + scale_fill_brewer(type = 'qual', palette = 3) + xlab("Longitude") + ylab("Latitude") + ggtitle("GDP Difference Global Visual")


```


## Interactive Visual
```{r interactive, fig.width=10, echo=FALSE}
x_axis_format <- list(
  title = "COUNTRY (HOVER OVER BARS FOR DETAILS)",
  showticklabels = FALSE
 
)

y_axis_format <- list(
  title = "GDP DIFFERENCE IN 2090"
)

plot_ly(combined, type="bar", x = ~name.x, y = ~diff_2090, color = ~value,
        size = ~value, text = ~paste("Internet Users per 100: ", value)) %>% layout(xaxis = x_axis_format, yaxis = y_axis_format)

```



